---
title: "ST537-G6-HW1"
author: "Group6"
date: "4/6/2020"
output: 
  pdf_document: 
    highlight: tango
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Intro


## Facets
This is placeholder for answers to facet questions
```{r facets, echo=TRUE, include=TRUE}
 ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy)) +
   
    facet_grid(drv ~ cyl)

 ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy)) +
    facet_wrap(. ~ class, nrow=3)

```

## Geoms

## Stats

## Positioning



## Coordinate Systems

Coordinate systems provide the ability to specify any given point in a plot. The most commonly known is Cartesian, which is just the standard X,Y format we are all used to. Each X and pair points to a specific place on the chart. But there are other features in R that can also be used to specify location in more complex ways.

--coord_flip() swaps the X and Y axes.
--coord_quickmap() used for special mapping in plots, for example on a map of Oregon.
--coord_polar() 

#### Coordinate Systems exercises
1. Create a polar plot: see below
2. What does labs() do? It allows adding titles, subtitles, captions, tags, and labels to the charts. Title was added to the charts below this way.
3. What is the difference between coord_map and coord_quickmap? Coord_map attempts to represent geography on a spherical plane, whereas coord_quickmap assumes the plane is flat. Coord_quickmap is best for smaller geographical areas where distortion won't cause problems with the data representation.
4. What does this plot tell you about the relationship between city and highway mpg? Higher city mileage is associated with higher highway mileage, but as hwy mileage goes up, there is a slight lag in city mileage increase ( the overall slope of the data seems to tilt left of the neutral line)
---What is geom_abline? this puts a line of a particular slope on the graph. By default it puts a 1x1 slope on the graph.
---What is coord_fixed? This enforces an "aspect ratio" so that the scales on the x and y axes are fixed in a ratio. 

```{r coordinate systems echo=TRUE, include=TRUE}

# Playing with bar and polar charts
bar <- ggplot(data = mtcars) +
  geom_bar(
    mapping = aes(x=cyl, fill=cyl),
    show.legend = FALSE,
    width = 1
  ) +
  theme(aspect.ratio = 1) 

bar + coord_flip() + labs(title = "No of Models by Cylinders")
bar + coord_polar() + labs(title = "No of Models by Cylinders ")
bar + labs(title="No of Models by Cylinders")

# Trying out abline and fixed

ggplot(data=mpg, mapping = aes(x = cty, y=hwy)) +
  geom_point() +
  geom_abline() +
  coord_fixed()
  
```
Hi
